Inferring neural networks from fluorescent calcium imaging using partial correlation and sample weighting

### 2014-05-30 article.tex
Present (partial) correlation from an independence perspective

% introduce the notion of dependence / independence
% TODO add a more formal definition of independence.
From a probabilistic point of view, the inference of the undirected network
can be formulated as inferring conditional dependences and
independences. Each neuron signal $i$ is a random continuous variable $X_i$,
which follows a probability density function $p_\mathcal{X_i}(x_i)$.
Two neurons $X_i$, $X_j$ will be said conditionally independent
$X_i \perp X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}$ to all
others neuron if
\begin{align*}
&p_{X_i, X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
    (x_i, x_j | x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\}) \\
=&p_{X_i | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
    (x_i | x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\})
p_{X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
    (x_j | x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\})
\end{align*}
for all real numbers $x_1,\ldots,x_p $ such that
$p_{X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
 (x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\}) > 0$.
Thus if $X_i \not\perp X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}$,
then we have the edge $i-j$ in the undirected graph.

% Introduce correlation => Estimator of I(X; Y)
%                       => Issue indirect effect
% Introduce partial correlation => Estimator of I(X; Y | Z)
%                               => Symmetric measure
% Introduce a custom solution to make the Y matrix assymetric
% Introduce Filtering

>>>>

% introduce the notion of dependence / independence
From a probabilistic point of view, the inference of the undirected network
can be formulated as inferring conditional dependences and
independences.  Even if \textit{independence
does no imply causation}[todo cite], it gives insight on the underlying connectivity
network. Each neuron signal $i$ is a random continuous variable $X_i$,
which follows a probability density function $p_\mathcal{X_i}(x_i)$.
Two neurons $X_i$, $X_j$ will be said conditionally independent
to a set of neurons $X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}$, denoted
by  $X_i \perp X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}$,  if
\begin{align*}
&p_{X_i, X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
    (x_i, x_j | x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\}) \\
=&p_{X_i | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
    (x_i | x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\})
p_{X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
    (x_j | x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\})
\end{align*}
for all real numbers $x_1,\ldots,x_p $ such that
$p_{X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}}
 (x_1,\ldots,x_p \backslash \left\{x_i, x_j\right\}) > 0$.

Under the hypothesis that the neurons follow a jointly Gaussian distribution
$p_{X_1,\ldots,X_p}(x_1,\ldots,x_p) \sim \mathcal{N}(\mu; \Sigma)$
of mean vector $\mu$ and of covariance matrix $\Sigma$, two neurons $X_i$
and $X_j$ are independent $X_i \perp X_j$ if they have zero Pearson correlation
coefficient
\[
\rho_{X_iX_j} = \frac{\Sigma_{ij}}{\sqrt{\Sigma_{ii} \Sigma_{jj}}}
\]
and are conditionally independent to all others neurons
$X_i \perp X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}$ if they
have zero partial correlation coefficient
\[
\rho_{X_i, X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}} =
\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}}.
\]

% Correlation => Issue indirect effect
% Partial correlation => better estimate of the graph  => Symmetric measure

Thus if $X_i \not\perp X_j | X_1,\ldots,X_p \backslash \left\{X_i, X_j\right\}$,
then we have the edge $i-j$ in the undirected graph.

% Introduce a custom solution to make the Y matrix asymmetric
% Introduce Filtering
